---
title: "Tech_val"
author: "Dylan Philippe & Alexander Lehner"
date: "10/04/2024"
output: html_document
---



Clear workplace and memory
```{r}
rm(list = ls())
gc()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "SET WORKING DIRECTORY HERE") # Set the knit root.dir here
```



```{r, echo=F}
# We import the necessary packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(raster, tmap, dplyr, sf, R.utils, ggplot2, rgeos, 
               nngeo, exactextractr, gifski, rlist, ncdf4, tidyr, zoo, 
               fasterize, minpack.lm)

options(scipen = 999) # This option allows to remove the scientific notation

# Deactivate the spherical geometry S2
sf::sf_use_s2(FALSE)

CRSproject <- "EPSG:4326"
EXTproject <- c(-25, 40, 35, 70) 
```


Load data 
```{r}
load("./processed_data/grid_europe.Rdata")
```


# Reference evapo
In the next chunks, we will compare our measure of reference evapotranspiration to the one computed in recently released datasets 
We will start by the one in the Version 3 of the Global Aridity Index and Potential Evapotranspiration Database by Zomer et al. (2022).
The dataset provide a measure of ET0 using climate data averaged over the 1970 - 2000 period (annual average).

```{r}
tech_val <- grid_europe
# Load the dataset from zomer et al. 2022
ET0_zomer <- raster("./inputs/reference_evapo/Global-AI_ET0_v3_annual/et0_v3_yr.tif")

# We crop it to the extend of the study and extract at grid level
ET0_zomer_eur <- crop(ET0_zomer, grid_europe)

tm_shape(ET0_zomer_eur) + tm_raster()

tech_val$ET0_zomer <- exact_extract(ET0_zomer_eur, tech_val, fun = "mean")

# The reference is in mm per year, we can convert it daily
tech_val <- tech_val %>% mutate(ET0_zomer = ET0_zomer/365)

# We load our reference evapotranspiration measure 
load("./processed_data/grid_evapo.Rdata")

# We select the year from 1970 - 2000
evapo_30 <- Reference_Evapo_grid[c(472:501)]

evapo_30_avg <- evapo_30[[1]]
# We repeatedely left join 

for (i in 2:30){
  evapo_30_avg <- left_join(evapo_30_avg, evapo_30[[i]] %>%
                                   st_drop_geometry() %>% dplyr::select(gid, ET0), by = "gid")
}


# We remove the gid and geometry variable and apply the rowmean
rows_evapo <- evapo_30_avg %>% st_drop_geometry() %>%
  dplyr::select(-c(gid))
rows_evapo$evapo_30 <- rowMeans(rows_evapo, na.rm = T)

tech_val$evapo_30 <- rows_evapo$evapo_30
```


# Ramakutty et al. 2002

## Ramankutty data
In this chunk we import and process the crop suitability from Ramankutty et al.
```{r}
raman <- raster("./inputs/Ramankutty/croppasturesuit.nc", var = "crops")


# We crop to the extend of our study
raman_eur <- crop(raman, grid_europe)

tech_val$raman_suit <- exact_extract(raman_eur, tech_val, fun = "mean")
```


This chunk will compute the grid averaging over the period 1961-1990 to compare with ramakutty
```{r}
# We load the gridded dataset of agricultural suitability
load("./processed_data/grid_suit.Rdata")

# We need to remove the small cells that we have droped 

grid_suit_30 <- grid_suit[c(462:491)]


grid_suit_30_avg <- grid_suit_30[[1]]
# We repeatedely left join 

for (i in 2:30){
  grid_suit_30_avg <- left_join(grid_suit_30_avg, grid_suit_30[[i]] %>%
                                   st_drop_geometry() %>% dplyr::select(gid, suit), by = "gid")
}


rows_suit <- grid_suit_30_avg %>% st_drop_geometry() %>%
  dplyr::select(-c(gid, f_c, f_ai, f_gdd, f_ph))
rows_suit$suit_30 <- rowMeans(rows_suit, na.rm = T)


# We bring the average back
tech_val$suit_30_raman <- rows_suit$suit_30
```


# FAO
```{r}
# We average our suitability to match the one of FAO 
load("./processed_data/grid_FAO.Rdata")

grid_suit_FAO <- grid_suit[c(472:501)]


grid_suit_FAO_avg <- grid_suit_FAO[[1]]
# We repeatedly left join 

for (i in 2:30){
  grid_suit_FAO_avg <- left_join(grid_suit_FAO_avg, grid_suit_FAO[[i]] %>%
                                   st_drop_geometry() %>% dplyr::select(gid, suit), by = "gid")
}

# Column 6 to 35 are scrop over the 30 years period: 1971 - 2000
rows_suit <- grid_suit_FAO_avg %>% st_drop_geometry() %>%
  dplyr::select(-c(gid, f_gdd, f_ai, f_c, f_ph))
rows_suit$suit_30_FAO <- rowMeans(rows_suit, na.rm = T)



tech_val$suit_30_FAO <- rows_suit$suit_30_FAO

# Now the FAO
tech_val <- left_join(tech_val, grid_FAO %>% st_drop_geometry() %>% dplyr::select(gid, FAO_suit), by = "gid")
```


```{r}
# Save the dataset 
save(tech_val, file = "./processed_data/tech_val.Rdata", overwrite = T)
```
